Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-06-28 analysis/memo2_TAS_reviewers.Rmd
modifications to paper and memo
additional minor edits to TAS memo; removing old draft

from: Authors
to: The American Statistican
subject: Reply to Second Round of Reviewer Comments
date: "`r format(Sys.Date(), '%B %d, %Y')`"
memorandum: true
graphics: false
width: .3
bibliography: references.bib
---

This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and an anonymous reviewer.

@@

We have made the app available to the public as a R package that includes the underlying dataset of historical strike data. We hope this will help statistics instructors in the way the TAS Associate Editor has in mind. However, because of time constraints, we are simply unable to provide in the paper's Supplement a complete tutorial tailored for classroom instruction by teachers of Bayesian statistics. On an introduction to a Bayesian approach to estimating illegal strike bias (albeit without historical strike data), we recommend @Kadane2018a and @Kadane2018b.

@@

The paper now contains links to the app as well as the code we used to generate the simulation results. We intend section 3.2 of the paper (and related Supplementary Material) as instructions for using the app. Those links are identifying, which is why we omitted them from the last draft of the paper.

@@

We appreciate this suggestion, but we worry that making this point here in terms of Type I and Type II error might lead to net confusion. The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more Type I error (falsely inferring illegal strike bias), because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence. Type I errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not. And Type I errors are worse still as the judge assigns more relative weight to any inference from historical strike data as compared to other items of evidence.

@@

Subsetting the historical data strike does not necessarily reduce incompatibility between the historical strike data and the current trial. Rather, it depends on what one _assumes_ about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) affects incompatibility. We stress "assumes", because the true values of $b_{hist}$ and $b_{curr}$ are not directly observable in actual cases. Accordingly, for any _subset_ of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$. The current version of the software prototype does not let users subset the historical data by some trial-level characteristic (e.g., defendant race). However, in principle, it could be modified to add that feature.

@@

We have revised the Discussion accordingly.  See also our reply to other reviewer below.

@@

```{r}
#| label: hypodata
#| echo: false

>>>>

from: Authors
to: The American Statistican
subject: Reply to Second Round of Reviewer Comments
date: "`r format(Sys.Date(), '%B %d, %Y')`"
memorandum: true
graphics: false
width: .3
#bibliography: references.bib
---

This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and an anonymous reviewer (Reviewer 2).

@@

We are ready to make the app available to the public as a R package that includes a de-identified version of the dataset of real historical strike data. (The link to the Github repository for the app identifies one of the authors. Accordingly, to preserve the anonymity that TAS requires, we have not yet put that link in the paper itself.) We hope this will help statistics instructors in the way the TAS Associate Editor has in mind. However, because of time constraints, we are simply unable to provide in the paper's supplement a complete tutorial tailored for classroom instruction by teachers of Bayesian statistics.

<!--
On an introduction to a Bayesian approach to estimating illegal strike bias (albeit without historical strike data), we recommend @Kadane2018a and @Kadane2018b.-->

@@

We intended section 3.2 of the paper (and related Supplementary Material) as instructions for using the app. Only because they are identifying, we have thus far omitted hyperlinks to the Github repositories for the app (ready to be available as an R package) as well as the code we used to generate the simulation results. When the TAS editor indicates it is appropriate, we will gladly add those links to the paper manuscript.

@@

We appreciate this suggestion, but we worry that making this point here in terms of Type I and Type II error might lead to net confusion. The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more false positives (false inferences of illegal strike bias, because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence. Such errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not. And such errors are worse still as the judge assigns more relative weight to the inference from historical strike data as compared to other items of evidence.

@@

Subsetting the historical data strike does not necessarily reduce incompatibility between the historical strike data and the current trial. Rather, it depends on what one _assumes_ about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) affects incompatibility. We stress "assumes", because the true values of $b_{hist}$ and $b_{curr}$ are not directly observable in actual cases. Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$. The current version of the software app does not let users subset the historical data by some trial-level characteristic (e.g., defendant race, courthouse location). However, in principle, it could be modified to add that feature.

@@

We have revised the Discussion accordingly.  See also our reply to Reviewer 2 below.

@@

```{r, echo = FALSE}
